---
title: "Differential Abundance by Residue"
output: html_notebook
---

This R notebook processes raw ProtMap peptide data across various citrullination conditions (uncitrullinated, PAD2-citrullinated, or PAD4-citrullinated) for several antigens and displays the resultant data by residue across the length of each protein sequence. Absolute abundance as well as differential abundance between conditions is plotted and saved to file. Resultant plots are shown in Figure 1 of the manuscript. 

```{r, setup, include=FALSE}
#Change directory to location of the GitHub repo. 
knitr::opts_knit$set(root.dir = '/media/aag7319/WDRed/Curran-et-al-2022_Cit_AgProc')
getwd()

```


```{r}
###WORKSPACE, LIBRARIES#####
source('./functions/cit_corr_fn.r')
library(seqinr)
library(ggplot2)
library(ggprism)
library(mutoss)
load("./objects/cit-sites-all-10.rda")
load("./objects/protmap_dat.rda")
```

Running this notebook will save plots to the below subfolder in your root directory (currently './results/'). The 'cutoff' line of twofold is drawn onto plots. 
```{r}
SAVE_FIGS <-TRUE
savedir <- paste0(getwd(), "/results/")
if (!dir.exists(savedir) && SAVE_FIGS) {dir.create(savedir, showWarnings = FALSE)}
cutoff <- log(2, base = 2) # 100% or greater DiffEx
```

Generate differential expression data from raw abundance data through the 'process1' function. Use plot functions to plot results in various formats - either DiffEx or raw abundance data, with or without citrullinated residues labeled. Source code for functions is in './functions/cit_corr_fn.r'. 
```{r}
cutoff <- log(2, base = 2) # 100% or greater DiffEx
#Generate DiffEx matrices
Ags <- c('ra33', 'fiba', 'fibb', 'fibg', 'vim', 'pad4')
Ag_len <- c(353,866,491, 453, 466, 663)
diffEx <- vector(mode = "list", length = 6);  names(diffEx) <- Ags
for( a in 1:length(Ags)){
  dat <- protmap_dat[[a]]
  tmp <- matrix(data= NA, nrow = Ag_len[a], ncol = 3)
  colnames(tmp) <- c("pad2.native", "pad4.native", "pad4.pad2")
  
  n <- dat[[1]]; p2 <- dat[[2]]; 

  p2.n <- process1(p2,n)
 
  tmp[,1] <- p2.n

  if(a < 6){ #Don't run these is Ag is pad4
    p4 <- dat[[3]]
    p4.n <- process1(p4,n)
    p4.p2 <- process1(p4, p2)
    tmp[,2] <- p4.n
    tmp[,3] <- p4.p2
  }
  
  diffEx[[a]] <- tmp
  
}
rm(dat,tmp,n,p2,p4,p2.n,p4.n,p4.p2)
```


Plot abundances and save in designated save directory. 
```{r}
if(SAVE_FIGS){
  
  subdir <- paste0(savedir, "/Abundances/")
  if (!dir.exists(subdir)){dir.create(subdir)}
  
  conds <- c("native", "pad2", "pad4")
  cutoff <- 6 #1e6
  for(f in 1: length(Ags)){
  c <- conds[1] #native
  a <- Ags[f]
  filename <- paste0(a,"_",c,"_abd.png")
  fullname <- paste0(subdir,filename)
  png(file = fullname, width = 750, height = 270)
  
  p <- plot0(protmap_dat[[f]][[1]], cutoff, paste(a,c))
  print(p)
  dev.off()
  
  
  c <- conds[2] #pad2
  a <- Ags[f]
  filename <- paste0(a,"_",c,"_abd.png")
  fullname <- paste0(subdir,filename)
  png(file = fullname, width = 750, height = 270)
  
  p <- plot0(protmap_dat[[f]][[2]],cutoff, paste(a,c))
  print(p)
  dev.off()
  
  if(f < 6){ #if not plotting pad4
  c <- conds[3] #pad4
  a <- Ags[f]
  filename <- paste0(a,"_",c,"_abd.png")
  fullname <- paste0(subdir,filename)
  png(file = fullname, width = 750, height = 270)
  
  p <- plot0(protmap_dat[[f]][[3]],cutoff, paste(a,c))
  print(p)
  dev.off()
  }
  }
  
  
}
```
Plot abundances with Cit sites marked, and save in designated save directory. 
```{r}
if(SAVE_FIGS){
  
  subdir <- paste0(savedir, "/Abd_cit/")
  if (!dir.exists(subdir)){dir.create(subdir)}
  
  conds <- c("pad2", "pad4")
  cutoff <- 6 #1e6 unadj abundance
  
  Ags <- names(cit_sites_all[[1]])
  for(f in 1: length(Ags)){
    exg <- c(1:3,5,6)
    
    c <- conds[1] #p2.native
    a <- Ags[f]
    filename <- paste0(a,"_",c,"_abd.png")
    fullname <- paste0(subdir,filename)
    if(length( cit_sites_all[[1]][[f]] > 0)){
      png(file = fullname, width = 750, height = 270)
      p <- plot3(protmap_dat[[exg[f]]][[2]], cit_sites_all[[1]][[f]], NULL, paste(a,c))
      print(p)
      dev.off()
    }
    
    if(f < 5){
      c <- conds[2] #p4.native. Don't do if Pad4 is Ag
      a <- Ags[f]
      filename <- paste0(a,"_",c,"_abd.png")
      fullname <- paste0(subdir,filename)
      
      if(length( cit_sites_all[[2]][[f]] > 0)){
      png(file = fullname, width = 750, height = 270)
      p <- plot3(protmap_dat[[exg[f]]][[2]], cit_sites_all[[2]][[f]], NULL, paste(a,c))
      print(p)
      dev.off()
      }
      
    }
  }
  
  
}
```
Plot differential expression and save in deisgnated save directory. 
```{r}
if(SAVE_FIGS){
  
  subdir <- paste0(savedir, "/DiffEx/")
  if (!dir.exists(subdir)){dir.create(subdir)}
  
  conds <- c("pad2_native", "pad4_native", "pad4_pad2")
  cutoff <- log(2, base = 2) #50% diffEx
  Ags <- c('ra33', 'fiba', 'fibb', 'fibg', 'vim', 'pad4')
  for(f in 1: length(Ags)){
    c <- conds[1] #native
    a <- Ags[f]
    filename <- paste0(a,"_",c,"_abd.png")
    fullname <- paste0(subdir,filename)
    png(file = fullname, width = 750, height = 270)
    
    p <- plot1(diffEx[[f]][,1], cutoff, paste(a,c))
    print(p)
    dev.off()
    
    if(f < 6){
    c <- conds[2] #pad2
    a <- Ags[f]
    filename <- paste0(a,"_",c,"_abd.png")
    fullname <- paste0(subdir,filename)
    png(file = fullname, width = 750, height = 270)
    
    p <- plot1(diffEx[[f]][,2],cutoff, paste(a,c))
    print(p)
    dev.off()
    
     #if not plotting pad4
      c <- conds[3] #pad4
      a <- Ags[f]
      filename <- paste0(a,"_",c,"_abd.png")
      fullname <- paste0(subdir,filename)
      png(file = fullname, width = 750, height = 270)
      
      p <- plot1(diffEx[[f]][,3],cutoff, paste(a,c))
      print(p)
      dev.off()
    }
  }
  
  
}
```

Plot differential expression with Cit sites marked, and save in designated save directory. 
```{r}
if(SAVE_FIGS){
  
  subdir <- paste0(savedir, "/DiffEx_cit/")
  if (!dir.exists(subdir)){dir.create(subdir)}
  
  conds <- c("pad2_native", "pad4_native")
  cutoff <- log(2, base = 2) #100% diffEx
  
  Ags <- names(cit_sites_all[[1]])
  for(f in 1: length(Ags)){
    c <- conds[1] #p2.native
    a <- Ags[f]
    i <- c(1,2,3,5,6) #skip fib.g
    filename <- paste0(a,"_",c,"_abd.png")
    fullname <- paste0(subdir,filename)
    
    if(length( cit_sites_all[[1]][[f]] > 0)){
    png(file = fullname, width = 750, height = 270)
    
    p <- plot41(diffEx[[i[f]]][,1], cit_sites_all[[1]][[f]], paste(a,c))
    print(p)
    dev.off()
    
    }
    
    if(f < 5){
      c <- conds[2] #p4.native. Don't do if Pad4 is Ag
      a <- Ags[f]
      filename <- paste0(a,"_",c,"_abd.png")
      fullname <- paste0(subdir,filename)
      
      if(length( cit_sites_all[[2]][[f]] > 0)){
      png(file = fullname, width = 750, height = 270)
      p <- plot41(diffEx[[f]][,2], cit_sites_all[[2]][[f]], cutoff, paste(a,c))
      print(p)
      dev.off()
      }
  
    }
  }
  
  
}
```

