---
title: "CitCutoffSweep"
author: "Alexander Girgis"
date: '2023-01-24'
output: html_document
---

```{r setup, include=FALSE}
#Change directory to location of the GitHub repo. 
knitr::opts_knit$set(root.dir = '/media/aag7319/WDRed/Curran-et-al-2022_Cit_AgProc')
getwd()
library(RColorBrewer)
library(viridis)
library(strex)
library(seqinr)
library(ggplot2)
source('./functions/fn.load.r')
fn.load()
```

This first few (lengthy) code chunks read in raw ProtMap mass spec data and compiles both abunudance values by residue, and the percentage each residue was observed in citrullinated vs uncitrullinated form.

```{r echo=TRUE, message=FALSE, warning=FALSE}
###LOAD,FORMAT DATA####
run1 <- read.all.sheets("./SourceData/CatBSH_digested_RA33_results_042921.xlsx")
run2 <- read.all.sheets(paste0("./SourceData/CatBSH_digested_fibrinogen",
                               "_vimentin_PAD4_results_072821.xlsx"))
run3 <- read.all.sheets("./SourceData/05_2021_12_28.xlsx")

v.interest <- c("Sequence", "Modifications", "Master Protein Accessions", 
                "Positions in Master Proteins")

#Extract from run1
s.idx <- c(15:17, 24:26, 18:20, 27:29, 21:23, 30:32)
d.idx <- c(match(v.interest, colnames(run1[[2]])), s.idx )

dat1 <- run1[[2]][,d.idx ]
dat1 <- dat1[(dat1$`Master Protein Accessions` == "P22626"), ] #subset to RA33
dat1$Protein <- "RA33"; dat1$file <- "run1"

#Extract from run2
dat2 <- list(fibrinogen = NA, vimentin = NA, pad4 = NA)
#fibrinogen
s.idx <- c(19,22,25,28,20,23,26,29,21,24,27,30)
d.idx <- c(match(v.interest, colnames(run2[[2]])), s.idx )
dat2[[1]] <- run2[[2]][,d.idx]
p.idx <- c( grep("P02671", dat2[[1]]$`Master Protein Accessions` ), 
            grep("P02675", dat2[[1]]$`Master Protein Accessions` ),
            grep("P02679", dat2[[1]]$`Master Protein Accessions` ))

dat2[[1]] <- dat2[[1]][p.idx, ]
#vimentin 
s.idx <- c(19,22,25,28,20,23,26,29,21,24,27,30) - 1
d.idx <- c(match(v.interest, colnames(run2[[4]])), s.idx )
dat2[[2]] <- run2[[4]][,d.idx]
p.idx <- grep("P08670", dat2[[2]]$`Master Protein Accessions` )
dat2[[2]] <- dat2[[2]][p.idx, ]
#PAD4
s.idx <-c(19,21,23,25, 20,22,24,26)
d.idx <- c(match(v.interest, colnames(run2[[6]])), s.idx )
dat2[[3]] <- run2[[6]][,d.idx]
p.idx <- grep("Q9UM07", dat2[[3]]$`Master Protein Accessions` )
dat2[[3]] <- dat2[[3]][p.idx, ]

#Extract from run3
dat3 <- list(fibrinogen = NA, vimentin = NA, pad4 = NA)
colnames(run3[[2]]) <- run3[[2]][1,]; # run3[[2]][-1,]
colnames(run3[[2]])[19:26] <- colnames(run3[[1]])[18:25]
#fibrinogen
s.idx <- c(19:21)
d.idx <- c(match(v.interest, colnames(run3[[2]])), s.idx )
p.idx <- c( grep("P02671", run3[[2]]$`Master Protein Accessions` ), 
            grep("P02675", run3[[2]]$`Master Protein Accessions` ),
            grep("P02679", run3[[2]]$`Master Protein Accessions` ))
dat3[[1]]<- run3[[2]][p.idx,d.idx]
#vimentin
s.idx <- c(22:24)
d.idx <- c(match(v.interest, colnames(run3[[2]])), s.idx )
p.idx <- grep("P08670", run3[[2]]$`Master Protein Accessions` )
dat3[[2]] <- run3[[2]][p.idx,d.idx]
#pad4
s.idx <- c(25:26)
d.idx <- c(match(v.interest, colnames(run3[[2]])), s.idx )
p.idx <- grep("Q9UM07", run3[[2]]$`Master Protein Accessions` )
dat3[[3]] <- run3[[2]][p.idx,d.idx]
###CLEAN DATA####
#obtain master sequence starts and ends. 
tmp <- bookEnds(dat1$`Positions in Master Proteins`)
dat1$Start <- tmp[,1]; dat1$End <- tmp[,2]; rm(tmp)
acc <-  c(NULL, "P08670", "Q9UM07")
for (i in 1:3){
  
  if(i == 3){
    tmp2 <- bookEnds(dat2[[i]]$`Positions in Master Proteins`, "Q9UM07")
    tmp3 <- bookEnds(dat3[[i]]$`Positions in Master Proteins`,  "Q9UM07")
  }else if(i == 2){
    tmp2 <- bookEnds(dat2[[i]]$`Positions in Master Proteins`, "P08670")
    tmp3 <- bookEnds(dat3[[i]]$`Positions in Master Proteins`,  "P08670")
  }else {
    tmp2 <- bookEnds(dat2[[i]]$`Positions in Master Proteins`)
    tmp3 <- bookEnds(dat3[[i]]$`Positions in Master Proteins`) 
  }
  
  dat2[[i]]$Start <- tmp2[,1];  dat2[[i]]$End <- tmp2[,2]
  dat3[[i]]$Start <- tmp3[,1];  dat3[[i]]$End <- tmp3[,2]
  rm(tmp2,tmp3)
}
#ignore all modifications except citrullination. 
#label file of origin prior to conacatenation.
dat1 <- only.cit(dat1)
for (i in 1:3){
  dat2[[i]] <- only.cit(dat2[[i]])
  dat3[[i]] <- only.cit(dat3[[i]])
  dat2[[i]]$file = "run2"
  dat3[[i]]$file = "run3"
}

###SUBSET BY CONDITION, GET CIT PCT####
#RA33, the easy one. 
ra33.pcts <- list(native = positional.cit.pct(dat1, c(5:10), "ra33"), 
                  pad2 = positional.cit.pct(dat1, c(11:16), "ra33") ,
                  pad4 = positional.cit.pct(dat1, c(17:22), "ra33") )
#Fibrinogen
fib.acc <- c("P02671","P02675", "P02679") #alpha, beta, gamma accession #s
tmp <- list(native = NA, pad2 = NA, pad4 = NA)
fib.pcts <- list(alpha = tmp, beta = tmp, gamma = tmp)
fib.run2.pcts <- list(alpha = tmp, beta = tmp, gamma = tmp)
fib.run3.pcts <- list(alpha = tmp, beta = tmp, gamma = tmp)
#dat2: 4 replicates. dat3: no replicates. 
#dat.idx is interpersed locations of nat.dat2, nat.dat3, pad2.dat2, pad2.dat3 etc. 
dat.idx <- list(c(5:8), c(5), c(9:12), c(6), c(13:16), c(7))
for(c in 1:3){ #for each condition 
  for (s in 1:3){ #for each subunit 
    #subset data from dat2
    d2.idx <- which(dat2$fibrinogen$`Master Protein Accessions` == fib.acc[s])
    d2 <- dat2[[1]][d2.idx,]
    d2.pct <- positional.cit.pct(d2, dat.idx[[2*c - 1]], fib.acc[s])
    
    #subset data from dat3
    d3.idx <- which(dat3$fibrinogen$`Master Protein Accessions` == fib.acc[s])
    d3 <- dat3[[1]][d3.idx,]
    d3.pct <- positional.cit.pct(d2, dat.idx[[2*c]], fib.acc[s] )
    
    fib.run2.pcts[[s]][[c]]<- d2.pct 
    fib.run3.pcts[[s]][[c]]<- d3.pct
    fib.pcts[[s]][[c]]<- (d2.pct + d3.pct) / 2
  }}

#Vimentin
vim.pcts <- list(native = NA, pad2 = NA, pad4 = NA)
vim.run2.pcts <- list(native = NA, pad2 = NA, pad4 = NA)
vim.run3.pcts <- list(native = NA, pad2 = NA, pad4 = NA)
dat.idx <- list(c(5:8), c(5), c(9:12), c(6), c(13:16), c(7))

for(c in 1:3){
  d2.pct <- positional.cit.pct(dat2[[2]], dat.idx[[2*c - 1]], "vim")
  d3.pct <- positional.cit.pct(dat3[[2]], dat.idx[[2 * c]], "vim")
  vim.run2.pcts[[c]] <- d2.pct
  vim.run3.pcts[[c]] <- d3.pct
  vim.pcts[[c]] <- (d2.pct + d3.pct) / 2
}

#pad4 
pad4.pcts <- list(native = NA, cit = NA)
pad4.run2.pcts <- list(native = NA, cit = NA)
pad4.run3.pcts <- list(native = NA, cit = NA)
dat.idx <- list(c(5:8), c(5), c(9:12), c(6) )

for(c in 1:2){
  d2.pct <- positional.cit.pct(dat2[[3]], dat.idx[[2*c-1]], "pad4")
  d3.pct <- positional.cit.pct(dat3[[3]], dat.idx[[2*c]], "pad4")
  
  pad4.run2.pcts[[c]] <- d2.pct 
  pad4.run3.pcts[[c]] <- d3.pct
  
  pad4.pcts[[c]] <- (d2.pct + d3.pct) / 2
}


data <- list(fiba = fib.pcts[[1]], fibb = fib.pcts[[2]], 
             fibg = fib.pcts[[3]], vim = vim.pcts, ra33 = ra33.pcts,
             pad4 = pad4.pcts)

data.run2 <- list(fiba = fib.run2.pcts[[1]], fibb = fib.run2.pcts[[2]], 
                  fibg = fib.run2.pcts[[3]], vim = vim.run2.pcts,
                  pad4 = pad4.run2.pcts)

data.run3 <- list(fiba = fib.run3.pcts[[1]], fibb = fib.run3.pcts[[2]], 
                  fibg = fib.run3.pcts[[3]], vim = vim.run3.pcts,
                  pad4 = pad4.run3.pcts)

#Group data by native or pad condition, NOT averaging runs but just throwing them 
#all together. used in scatter by condition section. 
data.native <- as.data.frame(data[[5]][[1]])
data.native$Ag <- "ra33"
data.native$residue <- c(1:353)
data.native$run <- "Run1"

data.pad2 <- as.data.frame(data[[5]][[2]])
data.pad2$Ag <- "ra33"
data.pad2$residue <- c(1:353)
data.pad2$run <- "Run1"
data.pad2$cond <- "pad2"

data.pad4 <- as.data.frame(data[[5]][[3]])
data.pad4$Ag <- "ra33"
data.pad4$residue <- c(1:353)
data.pad4$run <- "Run1"
data.pad4$cond <- "pad4"
aglen <- c(866,491,453,466,663)
ags <- c("fib.a", "fib.b", "fib.g", "vim", "pad4")
for (i in 1:5){
  x <- as.data.frame.matrix(data.run2[[i]][[1]] )
  y <- as.data.frame.matrix(data.run3[[i]][[1]] )
  x$residue <-c(1:aglen[i]) ; x$Ag <- ags[i]; x$run <- "Run2"
  y$residue <- c(1:aglen[i]); y$Ag <- ags[i]; y$run <- "Run3"
  data.native <- rbind(data.native, x, y)
  
  x <- as.data.frame.matrix(data.run2[[i]][[2]] )
  y <- as.data.frame.matrix(data.run3[[i]][[2]] )
  x$residue <-c(1:aglen[i]) ; x$Ag <- ags[i]; x$run <- "Run2"
  y$residue <- c(1:aglen[i]); y$Ag <- ags[i]; y$run <- "Run3"
  x$cond <- "pad2"
  y$cond <- "pad2"
  data.pad2 <- rbind(data.pad2, x, y)
  
  if (i < 5){
    x <- as.data.frame.matrix(data.run2[[i]][[3]] )
    y <- as.data.frame.matrix(data.run3[[i]][[3]] )
    x$residue <-c(1:aglen[i]) ; x$Ag <- ags[i]; x$run <- "Run2"
    y$residue <- c(1:aglen[i]); y$Ag <- ags[i]; y$run <- "Run3"
    x$cond <- "pad4"
    y$cond <- "pad4"
    data.pad4<- rbind(data.pad4, x, y)
  }
}


rm(d2,d2.pct,d3,d3.pct,dat.idx,dat1,dat2,dat3,run1,run2,run3,tmp, 
   acc,c,d.idx,d2.idx,d3.idx,fib.acc,i,p.idx,v.interest)

```

After calculating both abundances and citrullination by residue, we will visualize the extent to which citrullination was observed in 'native' protmap conditions (ie without PAD enzymes). A small degree of citrullination was observed under these conditions. We seek to establish a threshold for analysis which removes most 'native' citrullination sites but preserves the majority of citrullination sites observed in PAD2/PAD4 experimental conditions. The final thresholds used in analysis are shown in red in below plots (10% citrullinated and 1e6 total abundance).

```{r}
#Create a scatter of all native conditions in both runs to get a sense of baseline
# as well as in pad conditions. 

SAVE_FIGS <- TRUE
PARENT_DIR <- c("cit_cutoff_sweep") #This will be a subdir of master/results/
savedir <- paste0(getwd(), "/results/", PARENT_DIR)
if (!dir.exists(savedir) && SAVE_FIGS) {dir.create(savedir, showWarnings = FALSE)}

#Scatter by condition, but visualize boundaries
cit_cutoff <- 10
abd_cutoff <- 1e6
if(SAVE_FIGS){
  subdir <- paste0(savedir, "/scatter_by_cond/")
  if (!dir.exists(subdir)){dir.create(subdir)}
  specifier <- "Non-Normalized abundances" #used in plot titles
  #Native 
  del.idx <- which(data.native[,1] == 0)
  tmp <- as.data.frame.matrix( data.native[-del.idx, ])
  Ag <- tmp$Ag
  
  
  filename <- paste0("native_scatter_bounds.png")
  fullname <- paste0(subdir,filename)
  png(file = fullname, width = 420, height = 300)
  
  p1 <- ggplot(tmp, aes(x = `Total Abd`, y = `Percent Citrullinated`, colour = factor(Ag))) + 
    ylab("% Citrullinated") + xlab("log(Abd)") + 
    labs(title =paste("Native", specifier)) + geom_hline(yintercept = cit_cutoff, color = "red", size = 1.25) +
    geom_vline(xintercept = abd_cutoff, color = "red", size = 1.25) + 
    geom_point( size = 3) +
    theme_classic() + theme(panel.grid.major = element_line(colour = "light grey")) + 
    scale_x_log10(limits = c(31622.78, 3.162278e+12))
  #scale_x_log10(limits = c(3.162278e-06,3.162278)) 
  print(p1)
  dev.off()
  
  #Pad2
  del.idx <- which(data.pad2[,1] == 0)
  tmp <- as.data.frame.matrix( data.pad2[-del.idx, ])
  Ag <- tmp$Ag
  
  filename <- paste0("pad2_scatter_bounds.png")
  fullname <- paste0(subdir,filename)
  png(file = fullname, width = 420, height = 300)
  
  
  p2 <- ggplot(tmp, aes(x = `Total Abd`, y = `Percent Citrullinated`, colour = factor(Ag))) + 
    ylab("% Citrullinated") + xlab("log(Abd)") + 
    labs(title =paste("Pad2", specifier)) + 
    geom_point( size = 3) + geom_hline(yintercept = cit_cutoff, color = "red", size = 1.25) +
    geom_vline(xintercept = abd_cutoff, color = "red", size = 1.25) + 
    theme_classic() + theme(panel.grid.major = element_line(colour = "light grey")) + 
    scale_x_log10(limits = c(31622.78, 3.162278e+12))
  #scale_x_log10(limits = c(3.162278e-06,3.162278)) 
  print(p2)
  dev.off()
  
  #Pad4 
  del.idx <- which(data.pad4[,1] == 0)
  tmp <- as.data.frame.matrix( data.pad4[-del.idx, ])
  Ag <- tmp$Ag
  
  
  filename <- paste0("pad4_scatter_bounds.png")
  fullname <- paste0(subdir,filename)
  png(file = fullname, width = 420, height = 300)
  
  p3 <- ggplot(tmp, aes(x = `Total Abd`, y = `Percent Citrullinated`, colour = factor(Ag))) + 
    ylab("% Citrullinated") + xlab("log(Abd)") + 
    labs(title =paste("Pad4", specifier)) + geom_hline(yintercept = cit_cutoff, color = "red", size = 1.25) +
    geom_vline(xintercept = abd_cutoff, color = "red", size = 1.25) + 
    geom_point( size = 3) +
    theme_classic() + theme(panel.grid.major = element_line(colour = "light grey")) + 
    scale_x_log10(limits = c(31622.78, 3.162278e+12))
  #scale_x_log10(limits = c(3.162278e-06,3.162278)) 
  print(p3)
  dev.off()
  
  
  print(p1)
  print(p2)
  print(p3)
}


```
Curves below are the percent of 'native' citrullines excluded from analysis and percent of 'pad condition' citrullines included in analysis with a given Abundance threshold, and sweeping the minimum citrullination threshold. We seek to find a set of threhsold conditions which maximize the blue line (native citrullines excluded) and also maximize the red line (pad citrullines kept). The final cutoff values decided in manuscript are 1e6 abundance and 10% citrullinated. 
```{r}
#Sweep among these values: 
cit_cuts <- c(5:40)
abd_cuts <- c(0.5e5, 0.75e5, 1e5,0.75e6,1e6,1.25e6,1.5e6,1.75e6)

cut_res <- array(data = NA, dim = c(length(cit_cuts), length(abd_cuts), 3), 
                 dimnames <- list(cit_cut = cit_cuts, abd_cut = abd_cuts, 
                                  metric = c("Native_Exc", "Native_Exc_noPad", "Cit_Kept")))

del.idx <- which(data.native[,1] == 0)
tmp <- as.data.frame.matrix( data.native[-del.idx, ])
Ag <- tmp$Ag
total_native_cit <- dim(tmp)[1]

for (i in 0:(length(cit_cuts) * length(abd_cuts) - 1)){
  c <- floor(i / length(abd_cuts)) + 1
  a <- i %% length(abd_cuts) + 1
  abd_cutoff <- abd_cuts[a]
  cit_cutoff <- cit_cuts[c]

  #inc. Pad4 in calculation
  del.idx <- which(data.native[,1] == 0)
  tmp <- as.data.frame.matrix( data.native[-del.idx, ])
  Ag <- tmp$Ag
  total_native_cit <- length(unique(paste0(tmp$Ag, tmp$residue)))
  criteria_met <- which(tmp[,1] > cit_cutoff)[which(tmp[,1] > cit_cutoff) %in% which(tmp[,2] > abd_cutoff)]
  filt_cits <- tmp[criteria_met,]
  filt_cits <- unique(paste0(filt_cits$Ag, filt_cits$residue))
  total_filt <- length(filt_cits)
  pct_excl <- (total_native_cit - total_filt) / total_native_cit
  pct_excl <- (total_native_cit - total_filt) / total_native_cit * 100
  cut_res[c,a,1] <- pct_excl

  #exclude pad4 from calculation
  del.idx <- which(data.native[,1] == 0)
  tmp <- as.data.frame.matrix( data.native[-del.idx, ])
  tmp <- tmp[-which(tmp$Ag == "pad4"),]
  Ag <- tmp$Ag
  total_native_cit <- length(unique(paste0(tmp$Ag, tmp$residue)))
  criteria_met <- which(tmp[,1] > cit_cutoff)[which(tmp[,1] > cit_cutoff) %in% which(tmp[,2] > abd_cutoff)]
  filt_cits <- tmp[criteria_met,]
  filt_cits <- unique(paste0(filt_cits$Ag, filt_cits$residue))
  total_filt <- length(filt_cits)
  pct_excl <- (total_native_cit - total_filt) / total_native_cit * 100
  cut_res[c,a,2] <- pct_excl

  #How many of the pad2/pad4 samples were kept? 
  pads.combo <- rbind(data.pad2, data.pad4)
  del.idx <- which(pads.combo[,1] == 0)
  tmp <- as.data.frame.matrix( pads.combo[-del.idx, ])
  tmp <- tmp[-which(tmp$Ag == "pad4"),]
  Ag <- tmp$Ag
  total_native_cit <- length(unique(paste0(tmp$Ag, tmp$residue)))
  criteria_met <- which(tmp[,1] > cit_cutoff)[which(tmp[,1] > cit_cutoff) %in% which(tmp[,2] > abd_cutoff)]
  filt_cits <- tmp[criteria_met,]
  filt_cits <- unique(paste0(filt_cits$Ag, filt_cits$residue))
  total_filt <- length(filt_cits)
  pct_excl <- (total_native_cit - total_filt) / total_native_cit * 100
  cut_res[c,a,3] <- 100 - pct_excl

  
  
  
}



abd_cut = 1e5
a = which(abd_cuts == abd_cut)
p1 <- ggplot(data = data.frame(x = c(5:40), y = cut_res[,a,2]), aes(x = x, y = y)) + 
  geom_line(col = "dark blue") + theme_classic() + labs(title = paste("Abd =", abd_cut)) + xlab("%cit") + 
  ylab("% Excl/Kept") + geom_line(col = "dark red", data = data.frame(x = c(5:40), y = cut_res[,a,3]))
p1 


abd_cut = 1e6
a = which(abd_cuts == abd_cut)
p2 <- ggplot(data = data.frame(x = c(5:40), y = cut_res[,a,2]), aes(x = x, y = y)) + 
  geom_line(col = "dark blue") + theme_classic() + labs(title = paste("Abd =", abd_cut)) + xlab("%cit") + 
  ylab("% Excl/Kept") + geom_line(col = "dark red", data = data.frame(x = c(5:40), y = cut_res[,a,3]))
p2 


abd_cut = 1.5e6
a = which(abd_cuts == abd_cut)
p3 <- ggplot(data = data.frame(x = c(5:40), y = cut_res[,a,2]), aes(x = x, y = y)) + 
  geom_line(col = "dark blue") + theme_classic() + labs(title = paste("Abd =", abd_cut)) + xlab("%cit") + 
  ylab("% Excl/Kept") + geom_line(col = "dark red", data = data.frame(x = c(5:40), y = cut_res[,a,3]))
p3 





```
